Identifying 15-minute neighbourhoods with complete caring potential

### 2024-10-04 paper/Manuscript.qmd
Updates including Lucias comments

```{r library-setup}
library(biscale)
library(cowplot)
library(magick)
library(dplyr)
library(flextable)
library(ftExtra)
library(glue)
library(here)
library(mapview)
library(pander)
library(scales)
library(sf)
library(tmap)
library(tmaptools)
library(gt)
```

```{r}
load(glue::glue(here::here(), "/data/care_dest.rda"))
load(glue::glue(here::here(), "/data/access_15_wide.RData"))
```

@@

@@

The "15-Minute Neighbourhood" concept has been adopted by leaders around the world as a way to promote human-scale cities. The concept emphasizes a transport and land-use mix that supports short-trips to daily necessities by foot or bicycle from ones' residence. This study adopts the 15-minute neighbourhood concept through the lens of "Mobility of Care", a framework that foregrounds travel to care destinations, travel done predominately by women, in contrast to the better-studied travel to employment and leisure. While the 15-minute neighbourhood concept is flexible enough to consider all destination types, gendered examinations are relatively lacking in the literature and no research to date has focused explicitly on care destinations. In this study, we address these gaps by identifying which areas in Hamilton, Canada have the potential to be ‘caring 15-minute neighbourhoods'. To do so, a database of care destinations was compiled and used to estimate both the number (cumulative opportunities indicator) and diversity of mix (entropy measure) of destinations reachable within a 15-minute walk from all residential parcels. Parcel-level results were classified into 'caring' (quantity of reachable opportunities) and 'complete' (diversity of care destinations) neighbourhood typologies of varying degrees using machine learning techniques (Self-Organizing Maps (SOM) and hierarchical clustering). Inequalities in the relationship between the identified caring typologies and social-economic indicators of who currently reside in the neighbourhoods were then explored. Our results suggest the majority of 'caring and complete' neighbourhoods are within the urban core where a higher proportion of lower-income and single-parent households reside. Further, certain neighbourhoods in suburban and rural communities are 'caring and somewhat complete', showing the potential to become 'complete' through possible urban policy intervention. Taken together, this study offers a theoretical bridge to connect Mobility of Care and the 15-minute neighbourhoods concept, a data-driven classification and clustering methodology to assess neighbourhoods on a 'caring and complete' continuum, and discussion of relevant gender mainstreaming policy interventions.

@@

The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities [@teixeiraClassifying15minuteCities2024; @teixeiraClassifying15minuteCities2024]. As introduced in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. It is in opposition to the forces that supported the formation of single-use neighbourhoods such as industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. The "15-Minute City" is a utopic ideal, it is a city where one that accommodates the ability for all to reach important destinations within a walkable, bikeable or transit supportive radius. This form of neighbourhood planning would allow individuals to reclaim time spent on car mobility, giving way to sustainable modes and prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].

Though the term 15-Minute City is relatively new, neighbourhood planning practice aiming to foster community and local travel to amenities is not. As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022]. These planning forms have also been extensively critiqued: as summarized in @talenSocialSciencePlanned2017, urban planners struggled to determine suitably universal design aspects such as neighborhood size, the type and quality of amenities, and local and regional connectivity. Social scientists also emphasized that top-down neighbourhood planning approaches overestimated the built environment's influence on social life [@talenSocialSciencePlanned2017]. Neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced by some, but from bottom-up perspective. How to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate. For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners who frequently use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].

@@

In the last decades, the need for more sustainable, healthier, and liveable cities have received signficant attention. Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning principles to ones that center proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021]. In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021]. In @moreno_introducing_2021 where the 15-Minute City concept was introduced, four universal dimensions are outlined: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, <!-- (residential areas, work, commercial areas, education centres, health facilities and other basic institutions)--> and digitalisation. However, the framework presented has been criticized within academic and planning circles [@guzmanProximityEnoughCritical2024]. Critiques highlight the concept falls short in addressing the  pre-existing structural forces and individual characteristics that drive inequalities, influencing who benefits or could benefit the most from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023]. A 15-Minute City for all people is an aspirational ideal, but how will existing mobility and accessibility inequalities be addressed. Without directed and context-specific solutions, this increasingly popular concept risks becoming a place-branding slogan [@pozoukidou15MinuteCityDecomposing2021].

`r word_comment("Lucia, do I correctly reflect the plans you reviewed?", " ")`
Reflecting the flexible and aspirational presentation of the 15-Minute City; cities that have adopted the concept have done so using a diverse range of definitions and tools along with indistinct universal approaches. An early adopter was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012. It aims to foster an inclusive city development based on prosperity, education, health, and equity over a 30 year horizon. It promotes neighbourhood self-sufficiency and connectivity to city centers and centres of employment. The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017], similarly taking on an "all populations" approach. Later, other cities also incorporated proximity-based goals into their plans, adopting similar neutral entity approaches. For instance, in the plan that later catapulted the 15-Minute City into public discourse: Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign that foregrounded six social functions that should be easily accessible from any location, namely: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022]. These ideas inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021]. From the review of worldwide practice literature in @teixeiraClassifying15minuteCities2024, the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap: the 15-Minute City is aspirational but _how do we get there and whom will benefit_?

@@

`r word_comment("I am only vaguely familar with this literature as of these months - how does this sound?", " ")`
In addition, classification algorithms have been useful tools in the identification of mobility and spatial typologies in relation to transportation planning. Machine learning approaches have often been deployed. For instance, @burkeGeospatialAnalysisFramework2022 uses spatially constrained multivariate clustering, a regionalisation technique, to develop urban form typologies related to the 15-Minute City. The k-means algorithm is another popular method; @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns. Combing classification techniques is also common; for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group US neighbourhoods by minimising dissimilarity in attributes before applying k-means for further classification. Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the decision tree algorithm to partition the data into interpretable classifications.

@@

Rather than focusing on _all_ destinations, it may valuable to examine those related to caring activities. This re-framing offers a feminist perspective on  _what_ urban functions matter, connecting well with the 15-Minute City concept. Caring activities, which meet the physical, psychological, and emotional needs of others, are among the most essential and fundamental activities in society [@ilo_care_2018]. Yet, they are the most unequal, undervalued, and even devalued activities worldwide. Conventionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012]. According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5]. This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018]. In terms of spatial and transportation planning, almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. From this research motivation, Sánchez de Madariaga coined the term Mobility of Care in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While an undercurrent of research had examined these unique household-serving trips over the decades, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.

The Mobility of Care concept also explicitly integrates inter-sectional equity considerations. In considering Mobility of Care within the 15-Minute City one could address a common criticism leveled at the 15-Minute City [@guzmanProximityEnoughCritical2024]. Perhaps unsurprisingly given the gendered division of care work discussed, women have consistently been found to complete more mobility of care trips than men. In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s. While this gap is significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). As a product of the masculinst bias in planning, transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility. In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].

@@

`r word_comment("I need to clean this up and firmly reiterate the goals", " ")`
Due to it's popularity, the 15-Minute Concept is moving city planning towards a more people-centered approach focusing on accessibility to basic needs. While an aspirational ideal, the concept has faced criticism as a inadequately providing a way forward to address structural inequalities.  In this context, we bridge the 15-Minute City with the Mobility of Care concept, to add significance to what destinations and associated trips matter for all, but especially along gendered and socio-economic lines. We apply the "Caring 15-Minute City" concept to measure what neighourhoods in the mid-size Canadian city of Hamilton have the potential to be "caring 15 minute neighbourhoods". Making use of the tools reviewed and applied in the transportation research, the objectives of the paper are completed as follows:

1)	The cumulative accessibility measure and the entropy measure are used to quantify both the level of access and the diversity of access to care destinations within a 15-minute walk.
2) Based on these accessibility and diversity outputs, "Completely Caring 15-Minute Neighbourhood" typologies are generated using the SOM algorithm.  
3)	A decision tree is then applied to these typologies and socio-economic variables of populations who reside in these neighbourhoods. This is done to delineate data-driven profiles on which groups have the potential to benefit from what Completely Caring 15-Minute Neighbourhood.

@@

The mathematical formulation of takes the following form:
$$
D_{i} = \frac{-\sum_{t} [S_i^{t} / \sum_{t}S^{t}_{i} \times ln(S_i^{t} / \sum_{t}S^{t}_{i} )]}{ln(n_t)}
$$

>>>>

```{r library-setup}
library(biscale)
library(cowplot)
library(magick)
library(dplyr)
library(flextable)
library(ftExtra)
library(glue)
library(here)
library(mapview)
library(pander)
library(scales)
library(sf)
library(tmap)
library(tmaptools)
library(gt)
#install.packages("gtExtras")
library(gtExtras)
```

```{r}
load(glue::glue(here::here(), "/data/care_dest.rda"))
load(glue::glue(here::here(), "/data/access_15_wide.RData"))
load(glue::glue(here::here(), "/data/df_DA.RData"))
```

@@

```{r creating boxplot, eval=FALSE}
summary_boxplots <- access_15_wide |> select(c(scluster, starts_with("norm_"), div_allcats)) |>
  mutate(scluster = case_when(scluster == 1 ~ "A+",
                              scluster == 2 ~ "A",
                              scluster == 3 ~ "A-",
                              scluster == 4 ~ "B-",
                              scluster == 5 ~ "B",
                              scluster == 6 ~ "B+",
                              scluster == 7 ~ "D"),
         scluster = factor(scluster, ordered = TRUE,
                           levels = c("A+", "A", "A-","B+","B","B-","D")))

# summary_boxplots_Aplus <- summary_boxplots |> filter(scluster == "A+") |> select(-c("scluster"))
# summary_boxplots_A <- summary_boxplots |> filter(scluster == "A") |> select(-c("scluster"))
# summary_boxplots_Aminus <- summary_boxplots |> filter(scluster == "A-") |> select(-c("scluster"))
# summary_boxplots_Bplus <- summary_boxplots |> filter(scluster == "B+") |> select(-c("scluster"))
# summary_boxplots_B <- summary_boxplots |> filter(scluster == "B") |> select(-c("scluster"))
# summary_boxplots_Bminus <- summary_boxplots |> filter(scluster == "B-") |> select(-c("scluster"))
# summary_boxplots_D <- summary_boxplots |> filter(scluster == "D") |> select(-c("scluster"))

summary_boxplots_tidy <- summary_boxplots |>
  select(c("scluster","div_allcats",
           "norm_tot_copp_isc_norm_dep", "norm_tot_copp_isc_norm_err",
           "norm_tot_copp_isc_norm_groc", "norm_tot_copp_isc_norm_health")) |>
  rename("1"="div_allcats",
         "4"="norm_tot_copp_isc_norm_dep", 
         "2"="norm_tot_copp_isc_norm_err",
         "5"="norm_tot_copp_isc_norm_groc", 
         "3"="norm_tot_copp_isc_norm_health") |>
  pivot_longer(!scluster, names_to="variables", values_to = "normalized")
```

```{r, eval=FALSE}
box_plot <- ggplot(summary_boxplots_tidy, 
       aes(x=variables, y=normalized, fill=scluster)) +
  geom_boxplot(outlier.shape = NA,
               linewidth = 0.1) +
  scale_fill_manual(values=c("A+" = "#1A9850",
                               "A" = "#91CF60", 
                               "A-" = "#D9EF8B", 
                               "B+" = "#FEE08B",
                               "B" = "#FDAE61",
                               "B-" = "#F46D43",
                               "D" = "#D73027"),
                    name = " ") +
    scale_x_discrete(labels = c("1" = "Category Totals",
                                "2" = "ERRAND TOTALS",
                                "3" = "HEALTH TOTALS",
                                "4" = "DEPENDENT TOTALS",
                                "5" = "GROCERY TOTALS")) +
  coord_flip()+
  theme_bw() +
  theme(axis.title.y=element_blank()) +
  ylab("Min-max normalized values") 
box_plot
ggsave(file= file.path(here("figures/box-plot.png")))
```

```{r creating-table2-pt1}
summary_df_ALL <- access_15_wide |> select(c(starts_with("total_"), div_allcats)) |>
  group_by() |>
  summarise(across(everything(), list(mean = mean, sd = sd))) %>%
 mutate_if(is.numeric, round, digits =1) |>
  mutate_all(as.character)
summary_df_ALL <- cbind(data.frame(scluster = "TOTAL"),summary_df_ALL)

summary_df <- access_15_wide |> select(c(scluster, starts_with("total_"), div_allcats)) |>
  mutate(scluster = case_when(scluster == 1 ~ "A+",
                              scluster == 2 ~ "A",
                              scluster == 3 ~ "A-",
                              scluster == 4 ~ "B-",
                              scluster == 5 ~ "B",
                              scluster == 6 ~ "B+",
                              scluster == 7 ~ "D")) |>
  group_by(scluster) |>
  summarise(across(everything(), list(mean = mean, sd = sd))) %>%
 mutate_if(is.numeric, round, digits =1) |>
  mutate_all(as.character)

merge_every_two_columns <- function(df, num_rows) {
  # Loop through every second column starting from column 2
  for (col in seq(2, ncol(df), by = 2)) {
    if (col < ncol(df)) {  # Ensure the column has a next one to merge with
      for (i in 1:num_rows) {
        df[i, col] <- paste0(df[i, col], " (", df[i, col + 1], ")")
      }
    }
  }
  return(df)
}

summary_df <- merge_every_two_columns(summary_df, 7) |>  select(-contains("sd"))
summary_df_ALL <- merge_every_two_columns(summary_df_ALL, 1)  |>  select(-contains("sd"))
summary_df <- rbind(summary_df, summary_df_ALL)

summary_df <- summary_df |>
  pivot_longer(!scluster) |>
  pivot_wider(names_from =scluster )

summary_df_ALL <- summary_df_ALL |>
  pivot_longer(!scluster) |>
  pivot_wider(names_from =scluster )

summary_df$name <- c("GROCERY TOTALS", "Convenience Store", "Grocer",
                     "DEPENDENT TOTALS", "Community or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre",
                     "HEALTH TOTALS","Dentist","Hospital or clinic","Pharmacy",
                     "ERRAND TOTALS","Bank or ATM","Library", "Post office",
                     "Category diversity")

summary_df <- summary_df[,c(1,3,2,4,6,5,7,8,9)]
```

```{r, echo=FALSE}
df_DA_nogeom <- df_DA |> st_drop_geometry()
df_DA_nogeom <- df_DA_nogeom[,30:81]

df_DA_nogeom <- df_DA_nogeom |> mutate(
  Income = `v_CA21_907: Median after-tax income of household in 2020 ($)`,
  Avg_numb_of_kids = `v_CA21_498: Average number of children in census families with children`,
  Not_in_labforce = `v_CA21_6504: Not in the labour force`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
            Unemployed = (`v_CA21_6501: Unemployed`)/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income` ,
            Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
            walk_towork =  `v_CA21_7647: Walked`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            bike_towork =  `v_CA21_7650: Bicycle` / (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            pt_towork = `v_CA21_7644: Public transit`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            car_towork =  `v_CA21_7635: Car, truck or van`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            Owner_hld_CHN = `v_CA21_4308: % in core housing need (57)` ,
            Hld_ownership = `v_CA21_4306: % of owner households with a mortgage (58)` ,
            Hld_sheltercosts = `v_CA21_4307: % of owner households spending 30% or more of its income on shelter costs (55)` ,
            Tenant_sheltercosts = `v_CA21_4315: % of tenant households spending 30% or more of its income on shelter costs (55)` ,
            Tenant_subsi_hld = `v_CA21_4314: % of tenant households in subsidized housing (61)` ,
            Tenant_hld_CHH =`v_CA21_4316: % in core housing need (57)` ,
            Low_or_no_deg = (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate`) / (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate` + `v_CA21_5874: Postsecondary certificate, diploma or degree`))
```
```{r, echo=FALSE}
cors_df_DA_nogeom <- cor(df_DA_nogeom, use = "complete.obs") |> as.data.frame()
#cors_df_DA_nogeom[cors_df_DA_nogeom <0.7 | cors_df_DA_nogeom == 1] <- ""

One_parent_family_inspect <- cors_df_DA_nogeom |> select(One_parent_families)
LICO_inspect <- cors_df_DA_nogeom |> select(LICO)
Income_inspect <- cors_df_DA_nogeom |> select(Income)
```

@@

The "15-Minute Neighbourhood" concept has been adopted by leaders around the world as a way to promote human-scale cities. The concept emphasizes a transport and land-use mix that supports short-trips to daily necessities by foot or bicycle from ones' residence. This study adopts the 15-minute neighbourhood concept through the lens of "Mobility of Care", a framework that foregrounds travel to care destinations, travel done predominately by women, in contrast to the better-studied travel to employment and leisure. While the 15-minute neighbourhood concept is flexible enough to consider all destination types, gendered examinations are relatively lacking in the literature and no research to date has focused explicitly on care destinations. In this study, we address these gaps by identifying which areas in Hamilton, Canada have the potential to be ‘caring 15-minute neighbourhoods'. To do so, a database of care destinations was compiled and used to estimate both the number (cumulative opportunities indicator) and diversity of mix (entropy measure) of destinations reachable within a 15-minute walk from all residential parcels. Parcel-level results were classified into 'caring' (quantity of reachable opportunities) and 'complete' (diversity of care destinations) neighbourhood typologies of varying degrees using machine learning techniques (Self-Organizing Maps (SOM) and hierarchical clustering). Inequalities in the relationship between the identified caring typologies and social-economic indicators of who currently reside in the neighbourhoods were then explored. Our results suggest the majority of 'caring and complete' neighbourhoods are within the urban core where a higher proportion of lower-income households reside. Contrawise, neighbourhoods that are 'not caring and not complete' are located in peripheral regions that tend to have higher household incomes. Further, certain neighbourhoods in suburban and rural communities are 'caring and somewhat complete', showing the potential to become 'complete' through possible urban policy intervention. Implications of neighbourhoods that have a high potential to be improved, and socio-demographic characteristics associated with income such as single-parent households and economic poverty are discussed. Taken together, this study offers a theoretical bridge to connect Mobility of Care and the 15-minute neighbourhoods concept, a data-driven classification and clustering methodology to assess neighbourhoods on a 'caring and complete' continuum, and discussion of implications relevant to gender mainstreaming within transportation planning.

@@

The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities [@teixeiraClassifying15minuteCities2024]. As introduced in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. It is in opposition to the forces that supported the formation of single-use neighbourhoods such as industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. The "15-Minute City" refers to a city where relevant destinations within a walkable (or bikeable or transit supportive) radius are reachable by all. This form of neighbourhood planning would allow individuals to reclaim time spent on car mobility, giving way to sustainable modes and prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].

Though the term 15-Minute City is relatively new, neighbourhood planning practice aiming to foster community and local travel to amenities is not. As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022]. Along different dimensions, these planning forms have been extensively critiqued [@talenSocialSciencePlanned2017]. In this respect, neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced by some, but from bottom-up perspective using co-creation tools and meaningful resident participation to achieve equitable and justice neighbourhood forms [@mahmoudCocreationPathwayUrban2021]. How to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate. For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners who frequently use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].

From another perspective, feminist geographers, urban planners and other research groups agree that urban structure influences travel, but city planning fails short when planning for a neutral identity. In contrast, cities ought to be planned according to the multiple identities of the inhabitants [@vacchelli_towards_2018; @urbandevelopmentviennaGenderMainstreamingUrban2013]. Researchers have demonstrated how mobility patterns differ according to gender [@lawWomenTransportNew1999; @cresswell_gendered_2008; @levy2013travel; @little_gender_1994; @tronto_toward_1990]. Gender and social norms and cultural factors play a role in how, where, with whom, what for, and how far we move. Women have higher activities related to care [@ilo_care_2018], no matter at what life stage they are [@garciaromanGenderDifferencesTime2022]. For all these reasons the spatial organization of activities has a different impact according to gender, the 15-Minute City concept cannot be gender blind.

@@

In the last decades, the need for more sustainable, healthier, and livable cities have received significant attention. Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning principles to ones that center proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021]. In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021]. In @moreno_introducing_2021 where the 15-Minute City concept was introduced, four universal dimensions are outlined: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, <!-- (residential areas, work, commercial areas, education centres, health facilities and other basic institutions)--> and digitalisation. However, the framework presented has been criticized within academic and planning circles [@guzmanProximityEnoughCritical2024]. Critiques highlight the concept falls short in addressing the  pre-existing structural forces and individual characteristics that drive inequalities, influencing who benefits or could benefit the most from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023]. A 15-Minute City for all people is an aspirational ideal, but how will existing mobility and accessibility inequalities be addressed. Without directed and context-specific solutions, this increasingly popular concept risks becoming a place-branding slogan [@pozoukidou15MinuteCityDecomposing2021].

Reflecting the flexible and aspirational presentation of the 15-Minute City; cities that have adopted the concept have done so using a diverse range of definitions and tools along with indistinct universal approaches. A trail blazer was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012. Adopted before the '15-Minute City' concept of today, it aims to foster an inclusive city development based on prosperity, education, health, and equity over a 30 year horizon. The plan aims to promote neighbourhood self-sufficiency and connectivity to city centers and centres of employment. The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017], similarly taking on an "all populations" approach. Later, other cities also incorporated proximity-based goals into their plans, adopting similar neutral entity approaches. For instance, in the plan that later catapulted the 15-Minute City into public discourse: Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign that foregrounded six social functions that should be easily accessible from any location, namely: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022]. These ideas inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021]. From the review of worldwide practice literature in @teixeiraClassifying15minuteCities2024, the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap: the 15-Minute City is aspirational but _how do we get there and whom will benefit_?

@@

`r word_comment("I am only vaguely familar with this literature as of these months - how does this sound?", " ")`
In addition, classification algorithms have been useful tools in the identification of mobility and spatial typologies in relation to transportation planning. Machine learning approaches have often been deployed. For instance, @burkeGeospatialAnalysisFramework2022 uses spatially constrained multivariate clustering, a regionalisation technique, to develop urban form typologies related to the 15-Minute City. The k-means algorithm is another popular method; @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns. Similarly, [@montero2023role; @montero2023applying] cluster travel behavior profile typologies representing different population segments and place of residence in Barcelona Metropolitan Area. Combing classification techniques is also common; for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group US neighbourhoods by minimizing dissimilarity in attributes before applying k-means for further classification. Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the decision tree algorithm to partition the data into interpretable classifications.

@@

Rather than focusing on _all_ destinations, it may valuable to examine those related to caring activities. This re-framing offers a feminist perspective on  _what_ urban functions matter, connecting well with the 15-Minute City concept. Caring activities, which meet the physical, psychological, and emotional needs of others, are among the most essential and fundamental activities in society [@ilo_care_2018]. Yet, they are the most unequal, undervalued, and even devalued activities worldwide. Conventionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012]. According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care worldwide, a gender gap that varies geographically [@RN5]. This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018]. In terms of spatial and transportation planning, almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. From this research motivation, Sánchez de Madariaga coined the term Mobility of Care in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While an undercurrent of research had examined these unique household-serving trips over the decades, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.

The Mobility of Care concept also explicitly integrates inter-sectional equity considerations. In considering Mobility of Care within the 15-Minute City one could address a common criticism leveled at the 15-Minute City [@guzmanProximityEnoughCritical2024]. Perhaps unsurprisingly given the gendered division of care work discussed, women have consistently been found to complete more mobility of care trips than men. In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s. While this gap is significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). As a product of the masculinst bias in planning, transport surveys and tools most of the times do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility. In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].

@@

The objectives of this research are two-fold. 
- First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure.
- Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_  through an empirical examination of the mid-cited city of Hamilton, Canada. These neighbourhoods are then interpreted on the continuum of having (and not having) the region-relative land-use and transportation infrastructure to support _completely caring_ access. Who currently resides in which neighbourhood is examined, and ways to further expand the care accessibility story are discussed.

<!-- In this context, we bridge the 15-Minute City with the Mobility of Care concept, to add significance to what destinations and associated trips matter for all, but especially along gendered and socio-economic lines. We apply the "Caring 15-Minute City" concept to measure what neighourhoods in the mid-size Canadian city of Hamilton have the potential to be "caring 15 minute neighbourhoods". Making use of the tools reviewed and applied in the transportation research, the objectives of the paper are completed as follows: -->

<!-- 1)	The cumulative accessibility measure and the entropy measure are used to quantify both the level of access and the diversity of access to care destinations within a 15-minute walk. -->
<!-- 2) Based on these accessibility and diversity outputs, "Completely Caring 15-Minute Neighbourhood" typologies are generated using the SOM algorithm.   -->
<!-- 3)	A decision tree is then applied to these typologies and socio-economic variables of populations who reside in these neighbourhoods. This is done to delineate data-driven profiles on which groups have the potential to benefit from what Completely Caring 15-Minute Neighbourhood.  -->

@@

The mathematical formulation takes the following form:
$$
D_{i} = \frac{-\sum_{t} [S_i^{t} / \sum_{t}S^{t}_{i} \times ln(S_i^{t} / \sum_{t}S^{t}_{i} )]}{ln(n_t)}
$$
